---
title: "MGT6203 NFL Regression Project"
author: "Joanna Rashid"
date: "4/3/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Team Members

- Christopher Hom
- Lili Teister
- Joanna Rashid
- Stephen Yu

# Required Packages
```{r}
#install.packages("devtools")
#devtools::install_github(repo = "maksimhorowitz/nflscrapR")
#install.packages("nflreadr")

#library(nflscrapR) !!Lili, I could not get this package to load!!
library(dplyr)
library(nflreadr)
library(tidyverse)
library(lubridate)
library(SmartEDA)
library(corrplot)
library(ggcorrplot)
```

## Load games dataset

The games dataset is the main dataset that will be used for analysis. It contains a row for each NFL games from 1999 to 2021, but only seasons from 2010 will be included in the analysis.

```{r}
games_raw <- read.csv("http://www.habitatring.com/games.csv") %>%
  filter(season > 2009)

head(games_raw)
```
First, various attributes were selected to use as potential predicting or response variables.
One of those variables is moneyline, a betting wager, which can be converted with the following formula into the win probability based on bets placed before the game

```{r}

moneyline_to_odds_for <- function(ml){
  ifelse(ml < 0, (0-ml) / (0-ml+100), 100/(ml+100))
}

```

```{r}
games_select <- games_raw %>%
  mutate(home_prob = moneyline_to_odds_for(home_moneyline)) %>%
  select(game_id, # identifier for the game
         season, # year of season start
         game_type, # regular season, playoff game, etc
         week, # week number during the season
         gameday, # date of game
         weekday, # day of the week of the game, usually Sunday, Monday, or Thursday
         gametime, # time of the start of the game, eastern time
         away_team, # abbreviation for the away team
         home_team, # abbreviation for the home team
         home_score, # final score for the home team
         away_score, # final score for the away team
         location, # Home if the game was played at the home team's stadium, Neutral if in a neutral location
         result, # home team score - away team score
         total, # total points scored 
         overtime, # indicator for if the game went into OT
         home_rest, # number of days of rest the home team had since last game
         away_rest, # number of days of rest the away team had since last game
         home_moneyline, # used to calculate win prob for home team
         home_prob, # estimated win probability for the home team 
         total_line, # expected total points
         div_game, # indicator for if the teams playing each other are in the same division. A more consequential game for standings.
         roof, # type of stadium roof, such as dome, closed, or outdoors.
         surface, # type of grass on the field
         temp, # outside temp if the stadium is outdoors 
         wind # speed of wind if the stadium is outdoors
         )


```

Additionally, more information can be added to supplement the existing game information. 
One feature to add is how many injured players a team had during a game. However, injuries need to be combined with some sort of data on player value, since it really only matters if important players are injured. The simplest data I could find was rosters, which contains a player's approximate value for that year. 

```{r}
injuries <- load_injuries(c(2010:2019)) %>%
  mutate(season = as.integer(season))

head(injuries)

rosters <- read.csv("https://raw.githubusercontent.com/leesharpe/nfldata/master/data/rosters.csv") %>%
  filter(season > 2009)


injuries_roster <- injuries %>%
  left_join(rosters, by=c("season"="season", "team"="team", "full_name"="full_name")) %>%
  select(season,
         week,
         team,
         full_name,  # injured player full name
         report_status, # player's status on the official injury report
         games, # from rosters, number of games played total during that season for that team
         starts, # from rosters, number of total starts during that season for that team
         years, # from rosters, number of years player has been in the NFL
         av # from rosters, the player's Approximate Value for that year as defined by the Pro Football Reference metric
         )
```

## EDA
```{r}
games_select$game_type <- as.factor(games_select$game_type)

games_select$gameday <- as.Date(games_select$gameday)

games_select$weekday <- as.factor(games_select$weekday)

games_select$location <- as.factor(games_select$location)

games_select$roof <- as.factor(games_select$roof)

games_select$surface <- as.factor(games_select$surface)

#adds a binary variable for home team win or lose for logistic regression
games_select$win_lose <- ifelse(games_select$result > 0, 1 , 0)

#rest difference (home rest - away rest)
games_select$rest_diff <- (games_select$home_rest - games_select$away_rest)
```

```{r}
ExpData(data=games_select,type=2, fun = c("mean", "median", "var"))
```

```{r}
#numeric vars

numeric_data <- games_select %>% select(result, week, away_score, home_score, overtime, home_rest, away_rest, temp, wind, home_moneyline, home_prob, total_line, win_lose, rest_diff)

ExpNumViz(numeric_data,target="result",scatter=FALSE,fname=NULL,col="blue")
```
```{r}
#Correlation for numeric vars

numeric_corr <- cor(numeric_data)
corrplot(numeric_corr)
```

```{r}
#factors vars

factor_data <- games_select %>% select(game_type, weekday, location, overtime, roof, surface, win_lose)

factor_data$overtime <- as.factor(factor_data$overtime)
factor_data$win_lose <- as.factor(factor_data$win_lose)

ExpCatViz(factor_data,target=NULL,col ="slateblue4",clim=10,margin=2)
```

```{r}
#Correlation for factor vars

model.matrix(~0+., data=factor_data) %>% 
  cor(use="pairwise.complete.obs") %>% 
  ggcorrplot(show.diag = F, type="lower", lab=TRUE, lab_size=2)

```

## Initial Regression

```{r}
#initial linear model 

lm_initial <- lm(result ~ game_type + week + weekday + location + overtime + rest_diff, data = games_select)
summary(lm_initial)

#regressing only location on result
lm_home <- lm(result ~ location, data = games_select)
summary(lm_home)

log_home <- glm(win_lose ~ location, data = games_select, family = binomial(link = 'logit'))
summary(log_home)
```